Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-03-27 Paper/paper.tex
Update text

\section{Hydrogen photoionization rate}

@@

While four published composite quasar UV SEDs
\citep{2002ApJ...565..773T, 2001AJ....122..549V, 2012ApJ...752..162S,
  2014ApJ...794...75S} agree with Equation~(\ref{eqn:sed}), the
composite of \citet{2004ApJ...615..135S}, which uses more than 100
quasars at $z<0.1$, is somewhat steeper, with a EUV slope of $-0.56$.
The quasars considered by \citet{2004ApJ...615..135S} are fainter
($M_i(z=2)>-26$) than those considered by \citet{2015MNRAS.449.4204L},
which suggest that faint quasars may have a steeper EUV spectrum.  We
therefore also consider an SED of the form
\begin{equation}
f_\nu\propto\begin{cases}
               \nu^{-0.61\pm 0.01} & \text{if}~\lambda\geq 912~\text{\AA},\\
               \nu^{-0.56\pm 0.61} & \text{if}~600~\text{\AA}<\lambda<912~\text{\AA}.\\                
               \end{cases}
\label{eqn:sed_faint}
\end{equation}
for qsos with $M_{1450}<-23$.

The rate of emission of photons of frequency $\nu$ by quasars can be
written as
\begin{equation}
\dot n_\nu = \int dM_\nu \phi(M_\nu) \frac{L_\nu(M_\nu)}{h\nu}.
\end{equation}
Here the integral is over a suitable range of magnitudes, the
monochromatic luminosity $L_\nu(M)$ is related to the (absolute AB)
magnitude by \citep{1983ApJ...266..713O}
\begin{equation}
M_\nu = -2.5\log_{10}L_\nu+51.60,
\end{equation}
and $\phi$ is the luminosity function from Equation~(\ref{eqn:dpl}).
The luminosity at 1~Ry is related to that at 1450~\AA\ by
Equation~(\ref{eqn:sed}) or (\ref{eqn:sed_faint})
\begin{equation}
  L_{912}=L_{1450}\left(\frac{\nu_{912}}{\nu_{1450}}\right)^{-0.61}=L_{1450}\left(\frac{912}{1450}\right)^{0.61},
\end{equation}
depending on the qso luminosity.  The corresponding physical volume
emissivity is 
\begin{equation}
\epsilon_\nu = \dot n_\nu h\nu (1+z)^3.
\label{eqn:epsilon}
\end{equation}
Figure~\ref{fig:e912} shows the resultant 1~Ry emissivity when the
luminosity function is integrated down to $M_{1450}=-20$ (right panel)
and when it is integrated down to $M_{1450}=-18$ (left panel).
Similar to the previous figures, the black open circles in this figure
show the result form individual redshift bins, while the black curve
shown the result from best-fit global model.  Yellow curves show a
random sample from the global posterior distribution.  For comparison,
various other published determinations of the 1~Ry emissivity of
quasars are also shown in both panels of Figure~\ref{fig:e912}.  The
blue curve in these panels shows the AGN 1~Ry emissivity model of
\citet{2012ApJ...746..125H}, which is very similar to the model of
\citet{2007ApJ...654..731H}.  The red symbols with error bars show the
determination from the fits of \citet{2015AA...578A..83G}.  The maroon
curve is the published fit to the 1~Ry emissivity evolution in the
fits of \citet{2017MNRAS.466.1160M}.  The green curve is the model of
\citet{2015ApJ...813L...8M}.  Note that the integration limits for
these other models are not homogeneous.  \citet{2017MNRAS.466.1160M}
integrate the luminosity function down to $M_{1450}=-19$, while
\citet{2015AA...578A..83G} integrate down to $M_{1450}=-18$.  \gk{Add
  discussion about Madau and Khaire's integration limit from Gabor's 8
  March email here.  Also discussion on McGreer's data from Gabor's 12
  March email.}

The 1~Ry emissivity increases rapidly from $z=0$ up to $z=2$ and then
stays roughly constant until up to $z=3.5$.  At higher redshifts, due
to the relatively steeper faint end slope of the luminosity function,
the integration limit makes a large difference in the resultant
emissivity.  For the conservative case, in which we integrate under
the luminosity function down to $M_{1450}=-20$, the emissivity drops
rapidly at $z>3.5$, and is only about 10--20\% higher than that in the
\citet{2012ApJ...746..125H} model at $z\sim 6$.  However, in the model
in which we integrate down to $M_{1450}=-18$, the emissivity is higher
by a factor of almost 10 than that in the \citet{2012ApJ...746..125H}
model at $z\sim 6$.  In this case, our derived emissivities are
consistent with the values published by \citet{2015AA...578A..83G} at
$z=4$--$6$, although they still fall short by a factor of $\sim 2$.
Our derived emissivities are significantly smaller at $z>3.5$ than the
model of \citet{2015ApJ...813L...8M}.

@@

@@

\section{Helium reionization}

@@

The resultant reionization histories of hydrogen and helium are shown
in Figure XX.  We also show the reionization histories from
\citep{2012ApJ...746..125H} and \citep{2015ApJ...813L...8M} for
comparison.  As expected from the emissivity result of the previous
section, quasars cannot single handedly reionize hydrogen.  However
they can reionise \HeII.  The Helium reionization completes at
redshift $z=3.5$--$3.7$ depending on the luminosity down to which you
integrate the quasar luminosity function.  This is consistent with the
Helium \lya opacity measurements of \citet{2016ApJ...825..144W}.

@@

>>>>

\section{Contribution of AGN to reionization}

\subsection{Hydrogen-ionizing emissivity}

@@

While four published composite quasar UV SEDs
\citep{2002ApJ...565..773T, 2001AJ....122..549V, 2012ApJ...752..162S,
  2014ApJ...794...75S} agree with Equation~(\ref{eqn:sed}), the
composite of \citet{2004ApJ...615..135S}, which uses more than 100
quasars at $z<0.1$, is somewhat steeper, with a EUV slope of $-0.56$.
The quasars considered by \citet{2004ApJ...615..135S} are fainter
($M_i(z=2)>-26$) than those considered by \citet{2015MNRAS.449.4204L},
which suggest that faint quasars may have a steeper EUV spectrum.  We
therefore assume an SED of the form
\begin{equation}
f_\nu\propto\begin{cases}
               \nu^{-0.61\pm 0.01} & \text{if}~\lambda\geq 912~\text{\AA},\\
               \nu^{-0.56\pm 0.61} & \text{if}~600~\text{\AA}<\lambda<912~\text{\AA}.\\                
               \end{cases}
\label{eqn:sed_faint}
\end{equation}
for AGN with $M_{1450}<-23$.

The rate of emission of photons of frequency $\nu$ by quasars can be
written as
\begin{equation}
\dot n_\nu = \int dM_\nu \phi(M_\nu) \frac{L_\nu(M_\nu)}{h\nu}.
\end{equation}
Here the integral is over a suitable range of magnitudes, the
monochromatic luminosity $L_\nu(M)$ is related to the (absolute AB)
magnitude by \citep{1983ApJ...266..713O}
\begin{equation}
M_\nu = -2.5\log_{10}L_\nu+51.60,
\end{equation}
and $\phi$ is the luminosity function from Equation~(\ref{eqn:dpl}).
The luminosity at 1~Ry is related to that at 1450~\AA\ by
Equation~(\ref{eqn:sed}) or (\ref{eqn:sed_faint})
\begin{equation}
  L_{912}=L_{1450}\left(\frac{\nu_{912}}{\nu_{1450}}\right)^{-0.61}=L_{1450}\left(\frac{912}{1450}\right)^{0.61},
\end{equation}
depending on the AGN luminosity.  The corresponding physical volume
emissivity is 
\begin{equation}
\epsilon_\nu = \dot n_\nu h\nu (1+z)^3.
\label{eqn:epsilon}
\end{equation}
The open and filled black circles with errorbars in
Figure~\ref{fig:e912} show the resultant 1~Ry emissivity when the
luminosity function is integrated down to $M_{1450}<-21$ (open
circles) and when it is integrated down to $M_{1450}<-18$ (filled
circles).  Redshift bins that were removed from analysis due to large
systematic errors are not shown.  Similar to the previous figures, the
black open circles in this figure show the result form individual
redshift bins, while the black curve shown the result from best-fit
global model.  Due to the problems with the global luminosity function
models discussed above, we avoid using them to describe the emissivity
evolution.  Instead, we describe the emissivity evolution by the
5-parameter functional form \citep{2012ApJ...746..125H}
\begin{equation}
  \epsilon_{912}=\epsilon_0(1+z)^a\frac{\exp(-bz)}{\exp(cz)+d}.
  \label{eqn:e912fit}
\end{equation}
We fit this to the points shown in Figure~\ref{fig:e912} assuming a
Gaussian likelihood in each bin.  The corresponding best-fit curves
and the corresponding one-sigma uncertainty is shown in red
($M_{1450}<-18$) and blue ($M_{1450}<-21$).  The best-fit form of
Equation~\ref{eqn:e912fit} is given by 
\begin{multline}
  \epsilon_{912}=(10^{24.49}\mathrm{erg\, s^{-1}\, Hz^{-1}\, cMpc^{-3}})(1+z)^{7.57}\\\times\frac{\exp(-1.78z)}{\exp(1.01z)+29.20}
\end{multline}
for $M_{1450}<-18$, and by 
\begin{multline}
  \epsilon_{912}=(10^{24.11}\mathrm{erg\, s^{-1}\, Hz^{-1}\, cMpc^{-3}})(1+z)^{6.64}\\\times\frac{\exp(-0.65z)}{\exp(1.82z)+20.45}
\end{multline}
for $M_{1450}<-21$.  The 1~Ry emissivity increases rapidly from $z=0$
up to $z=2$ and then stays roughly constant until up to $z=3.5$.  At
higher redshifts, due to the relatively steeper faint end slope of the
luminosity function, the integration limit makes a large difference in
the resultant emissivity.  For the conservative case, in which we
integrate under the luminosity function down to $M_{1450}=-20$, the
emissivity drops rapidly at $z>3.5$For comparison, various other
published determinations of the 1~Ry emissivity of quasars are also
shown in both panels of Figure~\ref{fig:e912}.  We now comment on the
comparison of our results with these measurements.
\citet{2009A&A...507..781S} combined data from the Hamburg/ESO survey
and the SDSS to measure a quasar luminosity function at $z=0$ in their
$B_J$ band to minimize host galaxy contribution.  We convert their
$B_J$ magnitudes (Vega system) to $M_{1450}$ (AB) as
\begin{equation}
  M_{1450,\mathrm{AB}}=M_{B_J, \mathrm{Vega}}+0.59.  
\end{equation}
In this magnitude system, the luminosity function reported by
\citet{2009A&A...507..781S} has a very faint break luminosity of
$M_*=-18.87$.  The faint-end slope is also steep ($\beta=-2$).  We
integrate this luminosity function down to our limiting magnitudes of
$M_{1450}=-18$ and $-21$.  The resultant values are shown by the open
and closed brown circles at $z=0$ in Figure~\ref{fig:e912}.  Our
emissivity values at $z=0$ lie in between the
\citet{2009A&A...507..781S} measurements.  The convergence in the
emissivity is slower in their measurements than in our models because
of the shallower faint-end slope at this redshift in our model
($\beta\sim -1.22$).  We discuss the 1~Ry emissivity and the
corresponding hydrogen photoionization rate at $z=0$ below.

The brown dashed curve in Figure~\ref{fig:e912} shows the AGN 1~Ry
emissivity model of \citet{2012ApJ...746..125H}, which based on the
bolometric luminosity function model of \citet{2007ApJ...654..731H}.
Their assumed emissivity agrees reasonably well with our
$M_{1450}<-21$ case.  The green dotten curve in Figure~\ref{fig:e912}
shows the model presented by \citet{2015ApJ...813L...8M}.  This model
disagrees with ours rather severely at almost all redshifts in the
range considered here.  Note that a direct comparison with
\citet{2012ApJ...746..125H} and \citet{2015ApJ...813L...8M} is
difficult because these models integrate the luminosity function to a
fixed fraction of the break luminosity $L_*$ instead of a fixed
absolute magnitude.  The chosen limite by \citet{2012ApJ...746..125H}
and \citet{2015ApJ...813L...8M} is $0.01L_*$.  Using a limit that
depends on the break luminosity can be problematic as different quasar
surveys used to develop these emissivity models often report
strikingly different break luminosity values.  Even when these samples
are homogeneized appropriately, have such an integration limit is only
justified if the faint-end slope of the luminosity function is
sufficiently shallow.

The grey dashed curve shows the emissivity model of
\citet{2017MNRAS.466.1160M}.  These authors integrate the luminosity
function down to $M_{1450}=-19$.  However, their model strongly
disagrees with our inference, potentially due to biases resulting from
combining binned luminosity function estimates from different sources
inhomogeneously.  The green curves in Figure~\ref{fig:e912} show the
luminosity function derived from a variability-selected sample
observed using the Sloan telescope and the Multiple Mirror Telescope
(MMT) by \citet{2013A&A...551A..29P}.  This sample of 1877 quasars
does not overlap with the variability-selected sample reported by the
BOSS survey \citep{2013ApJ...773...14R}.  \citet{2013A&A...551A..29P}
assum constant luminosity function slopes with a break at $z=2.2$ and
assume a pure luminosity evolution.  This leads to a discontinuity in
the emissivity evolution in their model at $z=2.2$.  We convert their
$g$-band magnitudes to $M_{1450}$ and integrate the resultant
luminosity function down to our two chosen integration limits.  The
result is shown in Figure~\ref{fig:e912} by the solid and dashed
orange curves, which correspond to $M_{1450}<-21$ and $-18$,
respectively.  For $M_{1450}<-21$, our models are in reasonable
agreement at $z<2.2$.  At other redshifts and for $M_{1450}<-21$, the
models disagree by 10--30\%.  Figure~\ref{fig:e912} also shows the
model presented by \citet{2016A&A...587A..41P}.  We consider their
PLE$+$LEDE model, in which the faint-end slope is assumed to stay
constant with redshift.  Although the emissivity evolution in this
model is continuous at $z=2.2$, it is not smooth.  This model agrees
with our $M_{1450}<-18$ result to a better degree relative to the
\citet{2013A&A...551A..29P} model.  We also compare our result with
the PLE$+$LEDE luminosity function model presented by
\citet{2017A&A...608A..64C} by reanalysing the
\citet{2016A&A...587A..41P} sample with an improved $K$-correction
implementation.  \citet{2017ApJ...847L..15O} recently fit a double
power law luminosity function model to a combined binned samples of
\citet{2016ApJ...833..222J}, \citet{2010AJ....139..906W}, and
\citet{2015ApJ...798...28K}, and an X-ray-selected AGN candidate with
photometric redshift $z\sim 6$ reported by
\citet{2018MNRAS.474.2904P}.  We show the resultant 1~Ry emissivity by
orange rectangle in Figure~\ref{fig:e912}.  The lower side of the
rectangle corresponds to the emissivity for the
\citet{2017ApJ...847L..15O} model excluding the
\citet{2018MNRAS.474.2904P} AGN.  When this AGN candidate is included,
the emissivity is higher; this is shown by the upper end of the
rectangle.  The open rectangle corresponds to $M_{1450}<-18$ and the
closed rectangle corresponds to $M_{1450}<-21$.

The red open cirles with error bars in Figure~\ref{fig:e912} show the
determination from the fits of \citet{2015AA...578A..83G}.  These
authors also choose a break luminosity-dependent integration limit of
$0.01L_*$, which corresponds to $M_{1450}\sim -18$ in their model.
The resultant emissivities are factors of 2 to 3 higher than our
determinations for $z=4$--$6$.  This is puzzling at first sight,
because the AGN candidates reported by \citet{2015AA...578A..83G} are
consistent with our luminosity function estimates obtained without
these candidates (see Appendix~\ref{sec:conv}).  Our luminosity
function are also somewhat steeper than those reported by
\citet{2015AA...578A..83G}.  The discrepancy is explained by the
difference in the two luminosity function models at intermediate
magnitudes.  We find that the data prefer a much brighter break
luminosity than that infered by \citet{2015AA...578A..83G}.  This
reduces our emissivities relative to their estimates.  We go into this
issue in further detail in Appendix~\ref{sec:conv}.

\citet{2018AJ....155..131M} recently extended their previous Stripe~82
AGN sample by 1.5 magnitude and fit a double power law luminosity
function model to a combined SDSS $+$ Stripe~82 $+$ CFHTLS sample at
$4.7 < z < 5.1$.  These authors fixed the bright-end slope to
$\alpha=-4$ and obtained a shallower faint-end slope ($\beta=-1.97$)
than our models at this redshift ($\beta\sim -2.3$).  Rescaling their
LF to our cosmology and assumed our SED obtain the emissivities shown
by the green circles in Figure~\ref{fig:e912}.  Again the open and
filled circles corresponds to our two integration limits.  Our
emissivity agrees with the measurement of \citet{2018AJ....155..131M}
for $M_{1450}<-21$ but is higher than their measurement for
$M_{1450}<-18$, which is not surprising due to the difference in the
luminosity function slopes.  The emissivity derived from the
luminosity function of \citet{2018PASJ...70S..34A} is shown in yellow
in Figure~\ref{fig:e912}.  These authors report 1668 AGN candidates at
$3.5<z<4.3$, of which 76 have spectroscopic redshifts.  Their reported
faint-end slope is very flat ($\beta=-1.3\pm 0.05$), inconsistent with
our determination at all redshifts.  Their inferred emissivity is
consistent with our estimate for $M_{1450}\sim -21$.  Their emissivity
does not change significantly with the integration limit thanks to the
shallow slope.  We also compare our emissivities with the
determinations reported by \citet{2012ApJ...755..169M} at $z\sim 3.2$
and $z\sim 4$.  Their emissivity in reasonably agreement with our
model for $M_{1450}<-21$ at $z\sim 3.2$.  But their $M_{1450}<-18$
emissivity disagrees with all other high-redshift estimates, possibly
indicating a systematic error.  Our emissivity estimates are in good
agreement with those of \citet{2018MNRAS.474.2904P}.  Mention
Bongiorno.

@@

\subsection{Hydrogen photoionization rate}

@@

\subsection{Helium reionization}

@@

The resultant reionization histories of hydrogen and helium are shown
in Figure XX.  We also show the reionization histories from
\citep{2012ApJ...746..125H} and \citep{2015ApJ...813L...8M} for
comparison.  As expected from the emissivity result of the previous
section, quasars cannot single handedly reionize hydrogen.  However
they can reionise \HeII.  The Helium reionization completes at
redshift $z=3.5$--$3.7$ depending on the luminosity down to which you
integrate the quasar luminosity function.  This is consistent with the
Helium \lya opacity measurements of \citet{2016ApJ...825..144W}.
Discuss Mitra et al. 2018 and Khaire and Srianand 2018.  See Gabor's
email of 14 March 13:18.

@@

Some next steps at $z>4$: 1. include extended objects at faint
magntiudes that could be impacted by host galaxy light, 2. take
follow-up spectroscopy to confirm a fration of these as high-z broad
line AGN, 3. carefully assess host galaxy contamination.  See Gabor's
email of 14 March 13:18.
